---
title: "Reproducible Research: Peer Assessment 1"
author: "Sudhakar Athuru"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
1. Load data form the .csv file

```{r echo=TRUE}
setwd("C:/Sudhakar/coursera/05_ReproducibleResearch")
actdata <- read.csv("./data/activity.csv")
head(actdata)
names(actdata)
summary(actdata)
```

2. process data to aggregate steps per day
```{r echo=TRUE}
steps_by_day <- aggregate(steps ~ date, actdata, sum)
head(steps_by_day)
```

## What is mean total number of steps taken per day?
1. The following histogram shows number of total steps taken per day
```{r echo=TRUE}
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
```

2. Compute and print mean and median steps per day:
```{r computemean, echo=TRUE}
smean <- as.integer(mean(steps_by_day$steps))
smedian <- as.integer(median(steps_by_day$steps))
```
Mean steps per day are `r smean`. Median steps per day are `r smedian`.

## What is the average daily activity pattern?
1. Time series plot of the 5-minute interval (x-axis)and the average number of steps taken, averaged across all days (y-axis)process data to aggregate steps per day

```{r echo=TRUE}
steps_by_interval <- aggregate(steps ~ interval, actdata, mean)
summary(steps_by_interval)
plot(steps_by_interval$interval,steps_by_interval$steps, type="l", xlab="Interval", ylab="Number of Steps",main="Average Number of Steps per Day by Interval")
```

2. Which 5-minute interval, on average across all the days in the dataset,
contains the maximum number of steps?

```{r echo=TRUE}
max_interval <- steps_by_interval[which.max(steps_by_interval$steps),1]
```

`r max_interval` is the 5-minute interval with maximum number of steps.


## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
