---
title: "Reproducible Research: Peer Assessment 1"
author: "Sudhakar Athuru"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
### read the data form the .csv file

```{r echo=TRUE}
setwd("C:/Sudhakar/coursera/05_ReproducibleResearch")
actdata <- read.csv("./data/activity.csv")
head(actdata)
names(actdata)
summary(actdata)
```

### process data
```{r echo=FALSE, results="hide"}
actAgg <- tapply(actdata$steps, actdata$date, FUN = sum, na.rm = TRUE)
print(actAgg)
```

### process data to aggregate steps per day
```{r echo=TRUE}
steps_by_day <- aggregate(steps ~ date, actdata, sum)
head(steps_by_day)
```

## What is mean total number of steps taken per day?
The following histogram shows number of total steps taken per day
```{r echo=TRUE}
hist(steps_by_day$steps, main = paste("Total Steps Each Day"), col="blue", xlab="Number of Steps")
```

Compute and print mean and median steps per day:
```{r computemean, echo=TRUE}
smean <- as.integer(mean(steps_by_day$steps))
smedian <- as.integer(median(steps_by_day$steps))
```
Mean steps per day are `r smean`. Median steps per day are `r smedian`.


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
